Efficient Handling of String-Number Conversion

### 2020-04-02 final-version/main.tex
formulation

For ease of presentation, we  use the following toy example \changed{$$\Phi\!=\! \{``0"x\!=\! x``0", \sti{x} \!=\! \sti{y} ,|y|\!>\!|x|, 1000\!<\!|y| \}$$}
to explain the main ideas behind our decision procedure.  To make our terminology explicit: $\Phi$ states that \changed{$``0"$ concatenated with $x$ is the same as $x$ with $``0"$, the numeric value of the string $x$ is equivalent to that of $y$, $y$ is longer than $x$, and $y$ is longer than $1000$ characters.  Notice that $\Phi$ is satisfiable. E.g., it has a model $x=``00"$ and $y=``0^{1002}"$}.

Our decision procedure has two steps: the first step consists in over-approximating the set of input constraints into a set that falls in the chain-free fragment~\cite{abdulla2019chain}, which is decidable. Observe that we could  over-approximate the input constraint into any decidable fragment, e.g. the acyclic fragment~\cite{abdulla2014string} or the straight-line fragment~\cite{chen2019decision}. Our choice of the chain-free fragment~\cite{abdulla2019chain} is only motivated by the fact that the chain-free fragment is the {\em largest} known decidable fragment for that class of string constraints. In our example, we over-approximate the formula $\Phi$ by converting \changed{$``0"x = x``0"$ to two formulae $\{x_1=``0"x, x_2=x``0"\}$  and replacing the constraint $\sti{x}=\sti{y}$ with $n_x = n_y \wedge (n_x=-1 \vee (n_x\neq -1 \wedge x\in [0-9]^*)) \wedge (n_y=-1 \vee (n_y\neq -1 \wedge y\in [0-9]^*))$)}. 
Observe that if the over-approximation is UNSAT then our decision procedure  declares that the original formula is also UNSAT and terminates. Surprisingly, despite its simplicity, our over-approximation procedure works very well in practice as shown by our experimental results  (in Section \ref{section:evaluation}). Coming back to the formula $\Phi$, the over-approximation module will  return SAT  in this case.

@@

\changed{Although this toy example is seemingly trivial, all the string constraint solvers we tried (including Z3, CVC4, and Z3Str3) cannot solve it with a long timeout (10 min).}

>>>>

For ease of presentation, we  use the following toy example \changed{$$\Phi\!=\! \{``0"x\!=\! x``0", \sti{x} \!=\! \sti{y} ,|y|\!>\!|x|, 1000\!<\!|y| \}$$}
to explain the main ideas behind our decision procedure.  To make our terminology explicit: $\Phi$ states that \changed{$``0"$ concatenated with $x$ is the same as $x$ with $``0"$, the numeric value of the string $x$ is equivalent to that of $y$, $y$ is longer than $x$, and $y$ is longer than $1000$ characters.  Notice that $\Phi$ is satisfiable. E.g., it has a model $x=``00"$ and $y=``0^{1002}"$}.
\changed{Although this toy example is seemingly trivial, all the string constraint solvers we tried (including Z3, CVC4, and Z3Str3) cannot solve it within 10 minutes. }

%Our decision procedure has two steps: 
\changed{
Our new decision procedure solves the example in few seconds.
%
It proceeds in two steps:} the first step consists in over-approximating the set of input constraints into a set that falls in the chain-free fragment~\cite{abdulla2019chain}, which is decidable. Observe that we could  over-approximate the input constraint into any decidable fragment, e.g. the acyclic fragment~\cite{abdulla2014string} or the straight-line fragment~\cite{chen2019decision}. Our choice of the chain-free fragment~\cite{abdulla2019chain} is only motivated by the fact that the chain-free fragment is the {\em largest} known decidable fragment for that class of string constraints. In our example, we over-approximate the formula $\Phi$ by converting \changed{$``0"x = x``0"$ to two formulae $\{x_1=``0"x, x_2=x``0"\}$  and replacing the constraint $\sti{x}=\sti{y}$ with $n_x = n_y \wedge (n_x=-1 \vee (n_x\neq -1 \wedge x\in [0-9]^*)) \wedge (n_y=-1 \vee (n_y\neq -1 \wedge y\in [0-9]^*))$)}. 
Observe that if the over-approximation is UNSAT then our decision procedure  declares that the original formula is also UNSAT and terminates. Surprisingly, despite its simplicity, our over-approximation procedure works very well in practice as shown by our experimental results  (in Section \ref{section:evaluation}). Coming back to the formula $\Phi$, the over-approximation module will  return SAT  in this case.

@@
